---
title: "Homework_6"
author: "Laura J. Costello"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 6 
#### I pulled down a real dataset from the iNaturalist project Pollination Interactions on Plants: https://www.inaturalist.org/projects/pollinator-interactions-on-plants-pip-of-the-ne-us
#### Each row is an observation of a Fritillary butterfly in the northeast
#### Each observation includes a date, butterfly taxon name, and the name of the plant taxon it was observed on (mandatory in this project)
#### I want to look at how many plant species each butterfly species has been observed associating with

```{r}
# Read data in to a data frame from CSV exported from iNat
getwd()
list.files()
library(tidyverse)
setwd("C:/Users/ljcos/OneDrive/Documents/UVM/Spring 2025/Computational Biology/Lab5")
real_data <- read.csv(file="real_data2-reduced.csv",
                      header=TRUE,
                      sep=",",
                      fileEncoding="UTF-8"
                      ) #nrows=400
glimpse(real_data)

#grab just the columns I need
less_data <- data.frame(real_data["id"],real_data["observed_on"],
                        real_data["scientific_name"],
                        real_data["plant_species_name"]) 
head(less_data)
glimpse(less_data)

# clean date column to keep just the year
less_data$observed_on[] <- gsub("\\d+/\\d+/", 
                                "", 
                                less_data$observed_on[])

head(less_data)
length(less_data[[1]])

# clean data to keep only rows with plant at the species level
good_data_mask <- grep(" ",less_data$plant_species_name) #if there's a space, assume it's a good genus species name. vector of good indices
less_data <- less_data[good_data_mask, ] #keep only the good rows
length(less_data[[1]])

# clean data to keep only rows with butterfly at the species level
good_data_mask <- grep(" ",less_data$scientific_name) #if there's a space, assume it's a good genus species name. vector of good indices
less_data <- less_data[good_data_mask, ] #keep only the good rows
length(less_data[[1]])
```

Ok, now we have data

```{r}
# Group by year and scientific name, and count the number of unique associated plants for each butterfly that year
dataByYear <- less_data %>%
  group_by(observed_on,scientific_name) %>%
  summarize(plant_taxon_count=length(unique(plant_species_name)),n=n())
head(dataByYear)
summary(dataByYear)

# Group by year and scientific name, and count the number of unique associated plants for each butterfly that year
dataBySpecies <- less_data %>%
  group_by(scientific_name) %>%
  summarize(plant_taxon_count=length(unique(plant_species_name)),n=n())

```


Does the number of plants a butterfly is associated with follow a normal distribution? let's try a histogram
```{r}
hist(dataBySpecies$n)

#### Answer: absolutely not

```

Ok, how about a negative binomial??

```{r}
library(MASS)
require("fitdistrplus")
x <- df$plant_taxon_count
print(x)
plotdist(x,histo=TRUE,demp=TRUE)
fit <- fitdistr(x,"negative binomial")
print(fit)
dnbinom(x,size=1.1200373,mu=5.9586053)


curve(
  dnbinom(x,size=1,mu=6),
#      add=TRUE,
      lwd=2,
      col="red",
  from=0,
  to=50
)
```
