---
title: "Homework_5"
author: "Laura J. Costello"
date: "2025-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
### Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
sequence <- 3:10
#print(sequence)
n_dims <- sample(sequence, 1)
print(n_dims)
```
### Create a vector of consecutive integers from 1 to n_dims^2
```{r}
consec_vector <- 1:n_dims^2
print(consec_vector)
```
### Use the sample function to randomly reshuffle these values.
```{r}
consec_vector <- sample(consec_vector,length(consec_vector))
print(consec_vector)
```
### create a square matrix with these elements and print out the matrix.
```{r}
sqr_matrix <- matrix(data=consec_vector,nrow=sqrt(length(consec_vector)))
print(sqr_matrix)
```
### find a function in r to transpose the matrix and print again showing change
```{r}
sqr_matrix <- t(sqr_matrix)
print(sqr_matrix)
```
### calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
print(sqr_matrix[1,])
#sum of row 1
sum(sqr_matrix[1,])
#mean of row 1
mean(sqr_matrix[1,])
```
### read about the eigen() function and use it on your matrix
```{r}
print(eigen(sqr_matrix))
eigen_result <- eigen(sqr_matrix)
```
### look carefully at the elements of $values and $vectors in the output. What kind of numbers are these?
These numbers are complex: number + imaginary number
### dig in with the typeof() function to figure out their type.
```{r}
print(typeof(eigen_result[1][1])) #function returns a list of 2 items
print(eigen_result[1]) #print the first list item
print(eigen_result[[1]]) #print the value of the first list element
print(eigen_result[[1]][1]) #print the first item in that structure
print(typeof(eigen_result[[1]][1]))
print(eigen_result[[1]][[1]])
print(typeof(eigen_result[[1]][[1]]))
```
### if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.
True


## Problem 2
Create a list with the following named elements: my_matrix, which is a 4 x 4 matrix filled with random uniform values, my_logical which is a 100-element vector of TRUE or FALSE values. (Do this efficiently by setting up a vector of random values and then applying an inequality to it.), my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
my_matrix <- matrix(runif(16),4)
#print(my_matrix)

rand_vals <- runif(100)
my_logical <- rand_vals > runif(1)
#print(my_logical)

my_letters <- sample(letters,26)
#print(my_letters)

my_list <- list(my_matrix,my_logical,my_letters)
print(my_list)
```

Then, complete the following steps:

Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
new_list <- list(my_list[[1]][2,2],my_list[[2]][2],my_list[[3]][2])
print(new_list)
```

use the typeof() function to confirm the underlying data types of each component in this list

```{r}
typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])
```
combine the underlying elements from the new list into a single atomic vector with the c() function.

```{r}
atom_vector <- c(new_list[[1]],new_list[[2]],new_list[[3]])
print(atom_vector)
```

what is the data type of this vector?

```{r}
typeof(atom_vector)
```
## Problem 3

Create a data frame with the two variables (= columns) and 26 cases (= rows) below:

```{r}
#call the first variable my_unis and fill it with 26 random uniform values from 0 to 10
#call the second variable my_letters and fill it with 26 capital letters in random order.
my_unis <- runif(26,0,10)
print(my_unis)
my_letters <- sample(letters,26)
d_frame <- data.frame(my_unis,my_letters)
print(d_frame)
```

for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

```{r}
d_frame[sample(1:26,4),1] <- NA
print(d_frame)
```

for the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
which(!complete.cases(d_frame[1]))
```

re-order the entire data frame to arrange the second variable in alphabetical order

```{r}
order(d_frame$my_letters)
d_frame <- d_frame[order(d_frame$my_letters), ]
head(d_frame)
```

calculate the column mean for the first variable.

```{r}
print(d_frame[complete.cases(d_frame[[1]]),])
clean <- d_frame[complete.cases(d_frame[[1]]),]
print(clean)
clean[[1]]
mean(clean[[1]])
```
